---
title: "Capstone Questions"
author: "Arthur Cheib"
date: "2023-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## Libraries
library(tidyverse)
library(here)
library(sf)
library(readxl)
library(scales)
library(janitor)
library(openxlsx)
library(leaflet)
capstone_dir <- here('2023', 'Capstone')
capstone_raw_data <- here(capstone_dir, 'raw-data')
capstone_clean_data <- here(capstone_dir, 'clean-data')

```

```{r}
## Data
commercial_data_ar <- read_csv(here(capstone_clean_data, '3-commercial_licenses_cleaned.csv'))

## Keeping only EN columns and orderering the time period
commercial_data_en <- commercial_data_ar %>% 
  select(c(1), contains(c('EN', 'date', 'code'))) %>% 
  select(1,10,2,11,3,13,5,12,4,6,7,8,9)

## Geo data
# Rivers
saudi_rivers <- st_read(here(capstone_dir, 'geo-data', 'sau_rivers', 'riverl_sau.shp'))

# Airports
saudi_airports <- st_read(here(capstone_dir, 'geo-data', 'sau_airports', 'airp_sau.shp'))

# Saudi adm-division1
saudi_adm1 <- st_read(here(capstone_dir, 'geo-data', 'sau_adm_division', 'SAU_adm1.shp'))

head(st_geometry(saudi_rivers))
head(st_geometry(saudi_airports))
head(st_geometry(saudi_adm1))

# Initialize a leaflet map
m <- leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) # Add default OpenStreetMap base layer

# Add the shapefile layers to the map
m <- m %>% 
  #addPolygons(data = saudi_adm1, color = "#444444", weight = 1, fillOpacity = 0.5)
  addPolylines(data = saudi_rivers, color = "blue", weight = 1)
  addCircleMarkers(data = saudi_airports, color = "red", radius = 2)


saudi_airports %>% 
  leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>% 
  addCircleMarkers(color = "red", radius = 2) %>% 
  addPolygons(data = saudi_adm1, weight = 1, fillOpacity = 0.5, fill = '#444444', dashArray = TRUE)

saudi_adm1$VARNAME_1
commercial_data_en$municipality_EN %>% 
  unique() %>% 
  sort()
```

## Question 1: which Municipalities Attract the Most Business Activity?

**Objective:** Understand the geographical distribution of businesses.
**Skills Practiced:** Data aggregation, visualization, and interpretation.
**Expected Outcome:** A ranked list of municipalities based on the number of businesses. Visualizations (bar/line charts) that show the number of businesses in each municipality.

```{r}
## Commercial licenses per region
licenses_by_region <- commercial_data_en %>%
  count(municipality_EN, sort = TRUE)

# Plotting (by region)
licenses_by_region %>%
  ggplot(aes(x = reorder(municipality_EN, n), y = n)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(limits = c(0, 200000),
                     breaks = seq(0, 200000, by = 200000/8)) +
  coord_flip() + 
  labs(title = "Business Activity by Municipality", 
       y = "Number of Businesses", 
       x = "Municipality",
       caption = "Source: Saudi Arabia Open Data (www.od.data.gov.sa)") +
  geom_text(aes(label = scales::comma(n)), 
            hjust = -0.5, fontface = "bold", size = 3) +
  theme_minimal()

## Commercial licenses per municipality
commercial_data_en %>%
  count(municipality_branch_EN, sort = TRUE)


  
```


## Question 2: within Top 5 Municipalities, Which Branches See the Highest Concentration of Businesses?

**Objective:** Delve deeper into the municipalities to understand branch-wise distribution.
**Skills Practiced:** conditional filtering, aggregation, and visualization.
**Expected Outcome:** identification of top branches within major municipalities. Branch-wise distribution visualizations for top municipalities.

```{r}
## Getting the top5 regions
top_five_municipalities <- licenses_by_region %>% 
  head(5) %>% 
  pull(municipality_EN)

## Filtering and breaking it down by branches
top5_per_municipality_branch <- commercial_data_en %>% 
  filter(municipality_EN %in% top_five_municipalities) %>% 
  group_by(municipality_EN, municipality_branch_EN) %>% 
  summarize(total_business = n_distinct(id)) %>% 
  arrange(municipality_EN, desc(total_business)) %>% 
  ungroup()

```

## Question 3: what Are the Most Common Business Categories and Subcategories Across Different Municipalities?

**Objective:** identify dominant business types in various regions.
**Skills Practiced:** categorization, data aggregation, and visualization.
**Expected Outcome:** a breakdown of business categories and subcategories for each municipality. Pie charts or donut charts representing the share of each business category and subcategory

```{r}

```

## Question 4: how Do Business Issuances and Renewals Vary Across Municipalities and Branches?

**Objective:** understand the dynamics of new business entries vs. existing businesses.
**Skills Practiced:** data filtering, time series analysis, and visualization.
**Expected Outcome:** a visualization (stacked bar chart/line chart) showing the number of business issuances and renewals for each municipality and branch over time.

```{r}

```

## Question 5: predicting the Future: Which Municipalities and Branches Are Likely to See a Surge in Renewals in the Coming Year?

**Objective:** use the expiration data to forecast future renewal demands.
**Skills Practiced:** date-time manipulation, linear regression, and data visualization.
**Expected Outcome:** a forecasted list of municipalities and branches that might see a high number of renewals in the upcoming year. A heatmap or calendar visualization indicating the months when most renewals are due.

```{r}

```

## Question 6: how Does the Proximity to Major Cities, Rivers, and Airports Influence Business Distribution Across Municipalities?

**Objective:** determine if there's a correlation between the number or type of businesses in a municipality and its proximity to key geographical and infrastructural landmarks
**Skills Practiced:** geospatial analysis, data visualization on maps, correlation analysis.
**Expected Outcome:** a map heatmap map highlighting areas with dense business activity + major cities, rivers and airports (infra)

```{r}

```

